EuPPollNet: A European database of plant-pollinator networks

### 2024-10-01 Manuscript/EuPPollNet_Ms.qmd
authors

---
title: "EuPPollNet: A European database of plant-pollinator networks"
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
        \captionsetup[table]{labelformat=empty}
        \usepackage{lineno}
        \linenumbers
bibliography: references.bib
csl: apa_6th.csl
latex-tinytex: false
link-citations: TRUE
linkcolor: RoyalBlue
urlcolor: RoyalBlue
---

@@

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; https://doi.org/10.3030/101003476), first, data was directly requested from members of the Safeguard project in May 2022. Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019]. This approach maximized the potential number of studies that could be incorporated in this database. The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'. To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).

@@

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations). The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet. Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%). All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans. The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions. Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they conduct a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%). Hymenoptera and Diptera contain the highest number of species comprising each approximately 1,000 species in the database. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

@@

For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and programmatically fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.

@@

@@

Europe hosts approximately 5,000 species of pollinators and 25,000 species of plants that benefit from animal pollination. EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species. The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids). Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees. Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**). The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`). The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level. Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}. Phylogenetic and taxonomic coverage of the bee genera at European level. The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale. Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization. The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of bee species per genus at the European level. Dark grey bars indicate the total number of species per genus at the European level."}

#Circular layour
ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = 0.065, size=2.9, fontface=1) +
scale_size(range = c(0,4.5), guide = 'none')  +
scale_fill_continuous(type = "viridis", direction=-1) + 
guides(fill=guide_legend(), size = guide_legend()) +
theme(legend.position = "bottom") +
new_scale_fill() +
geom_fruit(data = d1, geom=geom_bar,
                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
                    pwidth=0.13, 
                    orientation="y", 
                    stat="identity",
                    color="black",
                    offset = 0.148,
            axis.params=list(
                         axis = "x",
                         text.size  = 2.5,
                         hjust  = 1.2,
                         vjust = 0.5,
                         nbreak = 3,
                         fontface = 2
                     ),
         grid.params=list()) +
guides(fill = "none")+
scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
geom_fruit(geom=geom_bar,
                    mapping=aes(y=label, x=Spp_number_Europe),
                    pwidth=0.12, 
                    orientation="y", 
                    stat="identity",
                    color="black",
                    offset = 0,
            axis.params=list(
                         axis = "x",
                         text.size  = 2.5,
                         hjust  = -0.2,
                         vjust = 0.5,
                         nbreak = 3,
                         fontface = 2
                     ),
         grid.params=list()) +
annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
new_scale_fill() +
geom_point2(aes(subset=node==131), color='orange', size=3) +
geom_point2(aes(subset=node==84), color='orange', size=3) +
geom_point2(aes(subset=node==73), color='orange', size=3) +
geom_point2(aes(subset=node==74), color='orange', size=3) +
geom_point2(aes(subset=node==111), color='orange', size=3)+
geom_point2(aes(subset=node==91), color='orange', size=3) +
annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
annotate("text", x = 2.55, y = 28.5, label = "Apidae", size= 3)+
annotate("text", x = 2.25, y = 51.2, label = "Megachilidae", size= 3) +
coord_cartesian(clip = "off")+
theme(legend.margin=unit(1,"cm"),
      legend.spacing = unit(1, "cm"),
      plot.margin = margin(4,0,4,0), 
      legend.position = c(0.6,0),
      legend.direction = "horizontal")

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}. Phylogenetic and taxonomic coverage of the plant families at European level. The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale. Additionally, a gradient of colors ranging from yellow to dark purple aids in this visualization. The coverage of species recorded in EuPPollNet per genus at the European level is depicted with orange and light grey bars, representing the percentage of species included and not included in the database, respectively, out of the total number of flowering plant species per family at the European level. Dark grey bars indicate the total number of species per family at the European level on logarithmic scale.", fig.width=8, fig.height=8}

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}. Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database. The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera. The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region. Note that a single study can contribute to more than one habitat or bioclimatic region. Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems. The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day.", fig.width=4.5, fig.height=8, fig.align='center'}

@@

#Outputs calculated in script 7_3
#Load lm output connectance ~ latitude
connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")
#Load lm output nestedness ~ latitude
normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")
#Load values of connectance
min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
#Load values of nestedness
min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
#Load kendall corr from connectance and spp
connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
#Load kendall corr from nestedness and spp
normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
#Load z values
#VAZNULL
z_percent_vaznull = readRDS("../Data/Manuscript_info/z_percent_vaznull.rds")
no_diff_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)
#CURVEBALL
z_percent_curveball = readRDS("../Data/Manuscript_info/z_percent_curveball.rds")
no_diff_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)

@@

@@

Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**). Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**). The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested. The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested. Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks. (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ = `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ = `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**). Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes. Note that residual connectance and normalised nestedness showed a moderate negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ = `r round(resid_connectance_nestedness_corr_pval, 2)`).

>>>>

---
title: "EuPPollNet: A European database of plant-pollinator networks"
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
        \usepackage{caption}
        \captionsetup[figure]{labelformat=empty}
        \captionsetup[table]{labelformat=empty}
        \usepackage{lineno}
        \linenumbers
bibliography: references.bib
csl: apa_7th.csl
latex-tinytex: false
link-citations: TRUE
linkcolor: RoyalBlue
urlcolor: RoyalBlue
---

@@

The EuPPollNet database includes published and unpublished studies compiled initially by a wide number of researchers and institutions within the European continent as defined by the European Environment Agency [@stanners1995]. As this database is the result of one of the work packages of the European project Safeguard (Safeguarding European wild pollinators; <https://doi.org/10.3030/101003476>), first, data was directly requested from members of the Safeguard project in May 2022. Second, the request was extended to data owners outside of the project, with data collection concluding in August 2024. These other data owners were identified by direct communication with colleagues suggested by Safeguard members and by directly searching for studies on Google Scholar of under-represented regions within the database. While Google Scholar lacks reproducibility [@gusenbauer2020], it still remains the most comprehensive search engine to date [@gusenbauer2019]. This approach maximized the potential number of studies that could be incorporated in this database. The search strings used were '*plant-pollinator interactions*' and '*plant-pollinator networks*'. To maintain high quality standards that will support robust future ecological research, we only included studies that met the following criteria: 1) studies containing time- and geo-referenced records of plant-pollinator interactions; and 2) studies that quantify interactions by documenting the contact between a floral visitor - referred to as a 'pollinator' throughout the manuscript, even though pollination efficiency is not evaluated - and the reproductive structure of a specific sampled plant (i.e., phyto-centric networks).

@@

The database contains `r study_number` independent published and unpublished studies conducted during the time period `r min_year` - `r max_year` in `r country_number` different countries (**Figure 1a** and **Figure 1b**; see **Figure S1** for exact locations). The studies differ in sampling effort and methodology, and thus documenting sampling methods and sampling effort is an important feature of EuPPollNet. Most studies took place within a single flowering season (`r format(round(one_year, 2), nsmall = 2)`%), sampled a given location for an average of `r format(round(average_sampling_days, 2), nsmall=2)` days, and exclusively sampled diurnal plant-pollinator interactions, with `r  most_common_method[[1]]` being the most common sampling method (`r format(round(most_common_method[[3]], 2), nsmall = 2)`%). All the studies documented interactions with Hymenopterans (with `r format(round(all_hymnoptera_percent, 2), nsmall = 2)`% considering all Hymenopterans, `r format(round(bee_percent, 2), nsmall = 2)`% only wild bees and `r format(round(bumblee_percent, 2), nsmall = 2)`% only bumblebees), `r format(round(Diptera_percent_metaweb, 2), nsmall = 2)`% documented interactions with Dipterans (with `r format(round(all_diptera_percent, 2), nsmall = 2)`% considering all Dipterans, `r format(round(syrphid_percent, 2), nsmall = 2)`% only syrphids and `r format(round(other_diptera_percent, 2), nsmall = 2)`% recorded syrphids plus bombylids or tachinid flies), `r format(round(Lepidoptera_percent_metaweb, 2), nsmall = 2)`% with Lepidopterans, and `r format(round(Coleoptera_percent_metaweb, 2), nsmall = 2)`% with Coleopterans. The database includes a total of `r prettyNum(interaction_number, big.mark=",")` distinct interactions. Most of the pollinator species belong to the orders Hymenoptera, Diptera, Lepidoptera and Coleoptera (`r format(round(spp_percent_main_order, 2), nsmall = 2)`%), which account for nearly all interactions in this database (`r format(round(main_orders_percent, 2), nsmall = 2)`%). Species that belong to other orders (`r format(round(100-spp_percent_main_order, 2), nsmall = 2)`%) are not explored in this study as they account for a minor fraction of the total interactions (`r format(round(100-main_orders_percent, 2), nsmall = 2)`%). Hymenoptera and Diptera contain the highest number of species, with each comprising approximately 1,000 species in the database. However, the majority of plant-pollinator interactions are from Hymenoptera species (`r format(round(Hymenoptera_interactions, 2), nsmall = 2)`%; **Figure 1c**). Notably, the western honey bee, _Apis mellifera_, represents `r format(round(apis_percent, 2), nsmall = 2)`% of the total interaction records from the database and an average of `r format(round(apis_average_network, 2), nsmall = 2)`% of the total interactions per network.

@@

For plants: (i) we initially verified the exact matches against the GBIF species checklist; (ii) we selected unmatched cases and fixed orthographic errors; (iii) we retrieved again taxonomic information for those unmatched cases, evaluated accuracy of fuzzy matching and manually fixed records that are still not found; (iv) finally, we used the World Flora Taxonomic Backbone [@govaerts2021\; WFO, July 7, 2022] as the ultimate filter for taxonomic information as we used it to calculate the plant taxonomic coverage of our database.

@@

#Load number of species from checklists and extrapolated one
bee_potential_spp = readRDS("../Data/Manuscript_info/bee_potential_spp.RData")
syrphid_potential_spp = readRDS("../Data/Manuscript_info/syrphid_potential_spp.RData")
lepidoptera_potential_spp = readRDS("../Data/Manuscript_info/lepidoptera_potential_spp.RData")
extrapolated_other_spp = readRDS("../Data/Manuscript_info/other_taxa_potential_spp.RData")
extrapolated_other_spp = round(extrapolated_other_spp)
```

Europe hosts approximately over 5,000 species of pollinators, including `r prettyNum(bee_potential_spp, big.mark=",")` bee species `r prettyNum(syrphid_potential_spp, big.mark=",")` syrphid species, `r prettyNum(lepidoptera_potential_spp, big.mark=",")` butterflies species, and about 1,700 species from other taxonomic groups, as estimated based on EuPPollNet average taxonomic coverage. Additionally, there are around 25,000 species of plants that benefit from animal pollination when excluding non-biotic pollinated species. EuPPollNet contains a total of `r prettyNum(pollinator_spp_number, big.mark=",")` pollinators and `r prettyNum(plant_spp_number, big.mark=",")` plant species. The coverage of the main pollinator groups occurring in Europe is `r format(round(bee_coverage, 2), nsmall=2)`% for bees, `r format(round(syrphid_coverage, 2), nsmall=2)`% for syrphids and `r format(round(lepidoptera_coverage, 2), nsmall=2)`% for butterflies (see **Figure S2** for coverage at the family level for bees and butterflies, and at the subfamily level for syrphids). Bees (i.e., Anthophila) constitute `r format(round(bee_interactions_percent, 2), nsmall=2)`% of the interactions in EuPPollNet, and `r format(round(bee_interactions_non_apis_percent, 2), nsmall=2)`% of the interactions when excluding honey bees. Within the database, `r format(round(bee_genera_percent_coverage, 2), nsmall=2)`% of bee genera have at least one species with interaction records, and the average coverage of species at the bee genus level is `r format(round(average_coverage_bee_genera, 2), nsmall=2)`% (**Figure 2**). The presence or absence of interaction records for bees does not follow a phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_bees, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_bees, 2), nsmall=2)`). The database coverage of all flowering plant species occurring in Europe is `r format(round(plant_coverage, 2), nsmall=2)`% (**Figure 3**), with an average coverage of `r format(round(mean_plant_fam_cov, 2), nsmall=2)`% at the plant family level. Approximately, half of the plant families have at least one species with interaction records (`r format(round(percent_plant_fam_with_data, 2), nsmall=2)`%), and the presence or absence of interaction data for the different plant species also does not follow a statistically relevant phylogenetic pattern ($\lambda$ = `r format(round(lambda_phylosig_plants, 2), nsmall=2)`; _P_ = `r format(round(p_phylosig_plants, 2), nsmall=2)`).

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.width=10, fig.height=10, fig.cap= "\\textbf{Figure 2}. Phylogenetic and taxonomic coverage of bee genera at European level. The number of interactions recorded per genus in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple. Additionally, the coverage of species recorded in EuPPollNet for each genus is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database, out of the total number of bee species in Europe. Dark grey bars represent the total number of species per genus at European level."}

#Circular layour
ggtree(tree_family_interactions, size=0.6, open.angle=1, alpha=1)+ 
geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = 0.065, size=2.9, fontface=1) +
scale_size(range = c(0,4.5), guide = 'none')  +
scale_fill_continuous(type = "viridis", direction=-1) + 
guides(fill=guide_legend(), size = guide_legend()) +
theme(legend.position = "bottom") +
new_scale_fill() +
geom_fruit(data = d1, geom=geom_bar,
                    mapping=aes(y=Genus1, x=Percent, fill = fam_group),
                    pwidth=0.13, 
                    orientation="y", 
                    stat="identity",
                    color="black",
                    offset = 0.148,
            axis.params=list(
                         axis = "x",
                         text.size  = 2.5,
                         hjust  = 1.2,
                         vjust = 0.5,
                         nbreak = 3,
                         fontface = 2
                     ),
         grid.params=list()) +
guides(fill = "none")+
scale_fill_manual(values = c(a = "tan2", b = "lightgrey"), guide = "none")+
geom_fruit(geom=geom_bar,
                    mapping=aes(y=label, x=Spp_number_Europe),
                    pwidth=0.12, 
                    orientation="y", 
                    stat="identity",
                    color="black",
                    offset = 0,
            axis.params=list(
                         axis = "x",
                         text.size  = 2.5,
                         hjust  = -0.2,
                         vjust = 0.5,
                         nbreak = 3,
                         fontface = 2
                     ),
         grid.params=list()) +
annotate("text", x = 14.35, y = 70, label = "EuPPollNet spp\n coverage (%)", size= 3, fontface= "bold") +
annotate("text", x = 16, y = 70, label = "European\n species", size= 3, fontface= "bold")+
annotate("text", x = 12.2, y = 70, label = "EuPPollNet\n interactions", size= 3, fontface= "bold")+ 
new_scale_fill() +
geom_point2(aes(subset=node==131), color='orange', size=3) +
geom_point2(aes(subset=node==84), color='orange', size=3) +
geom_point2(aes(subset=node==73), color='orange', size=3) +
geom_point2(aes(subset=node==74), color='orange', size=3) +
geom_point2(aes(subset=node==111), color='orange', size=3)+
geom_point2(aes(subset=node==91), color='orange', size=3) +
annotate("text", x = 9.45, y = 2.7, label = "Melittidae", size= 3)+
annotate("text", x = 7.3, y = 6.4, label = "Andrenidae", size= 3)+
annotate("text", x = 10.4, y = 12.4, label = "Colletidae", size= 3)+
annotate("text", x = 5.45, y = 17.3, label = "Halictidae", size= 3)+
annotate("text", x = 2.25, y = 26.8, label = "Megachilidae", size= 3)+
annotate("text", x = 2.5, y = 51, label = "Apidae", size= 3) +
coord_cartesian(clip = "off")+
theme(legend.margin=unit(1,"cm"),
      legend.spacing = unit(1, "cm"),
      plot.margin = margin(4,0,4,0), 
      legend.position = c(0.6,0),
      legend.direction = "horizontal")

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}. Phylogenetic and taxonomic coverage of the plant families at European level. The number of interactions recorded per family in the database is illustrated using circles, with their sizes proportional to the number of interactions on a logarithmic scale, complemented by a gradient of colors ranging from yellow to dark purple. Additionally, the coverage of species recorded in EuPPollNet for each family is depicted using two types of bars: orange bars representing the percentage of species included in the database and light grey bars indicating the percentage of species not included in the database out of the total number of bee species per family in Europe. Dark grey bars represent the total number of species per family at European level.", fig.width=8, fig.height=8}

@@

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 5}. Proportion of species from the major pollinator orders by habitat types and bioclimatic regions in the EuPPollNet database. The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera. The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region. Note that a single study can contribute to more than one habitat or bioclimatic region. Areas with a greater number of studies are more likely to depict accurate proportions of the different pollinator orders in those systems. The Pannonian and Steppic bioregions were excluded from this visualization because they contain only few networks from a single sampling day and site.", fig.width=4.5, fig.height=8, fig.align='center'}

@@

#Outputs calculated in script 7_3
#Load lm output connectance ~ latitude
connectance_lat_r2= readRDS("../Data/Manuscript_info/connectance_lat_r2.rds")
connectance_lat_pval = readRDS("../Data/Manuscript_info/connectance_lat_pval.rds")

connectance_lat_pval = ifelse(connectance_lat_pval <0.01, 0.01, connectance_lat_pval)

#Load lm output nestedness ~ latitude
normalised_nestedness_lat_r2 = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_r2.rds")
normalised_nestedness_lat_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_lat_pval.rds")

#Load values of connectance
min_connectance = readRDS("../Data/Manuscript_info/min_connectance.rds")
max_connectance = readRDS("../Data/Manuscript_info/max_connectance.rds")
mean_connectance = readRDS("../Data/Manuscript_info/mean_connectance.rds")
#Load values of nestedness
min_nodfc = readRDS("../Data/Manuscript_info/min_nodfc.rds")
max_nodfc = readRDS("../Data/Manuscript_info/max_nodfc.rds")
mean_nodfc = readRDS("../Data/Manuscript_info/mean_nodfc.rds")
#Load kendall corr from connectance and spp
connectance_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_tau.rds")
connectance_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/connectance_spp_mean_corr_pval.rds")
#Load kendall corr from nestedness and spp
normalised_nestedness_spp_mean_corr_tau = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_tau.rds")
normalised_nestedness_spp_mean_corr_pval = readRDS("../Data/Manuscript_info/normalised_nestedness_spp_mean_corr_pval.rds")
#Load z values
#VAZNULL
z_percent_vaznull = readRDS("../Data/Manuscript_info/z_percent_vaznull.rds")
no_diff_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_vaznull = z_percent_vaznull %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)
#CURVEBALL
z_percent_curveball = readRDS("../Data/Manuscript_info/z_percent_curveball.rds")
no_diff_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "No statistical difference") %>% 
pull(percent)
over_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "Over-represented") %>% 
pull(percent)
under_z_percent_curveball = z_percent_curveball %>% 
filter(infra_over_represented == "Under-represented") %>% 
pull(percent)

@@

#To avoid having values under 0.01
resid_connectance_nestedness_corr_pval = ifelse(resid_connectance_nestedness_corr_pval < 0.01, 0.01, resid_connectance_nestedness_corr_pval)

@@

Connectance values ranged between `r round(min_connectance, 2)` to `r round(max_connectance, 2)` ($\bar{x}$ = `r round(mean_connectance, 2)`) and followed a negative exponential relationship with the number of species per network (Kendall $\tau$ = `r round(connectance_spp_mean_corr_tau, 2)`, _P_ < 0.01; **Figure 6a**). Nestedness values (NODFc) ranged between `r round(min_nodfc, 2)` to `r round(max_nodfc, 2)` ($\bar{x}$ = `r round(mean_nodfc, 2)`), and were not independent of the mean number of species (Kendall $\tau$ = `r round(normalised_nestedness_spp_mean_corr_tau, 2)`, _P_ = `r round(normalised_nestedness_spp_mean_corr_pval, 2)`; **Figure S4**). Empirical networks did not show statistically different nestedness to the simulated ones (**Figure 6b** and **Figure S5**). The 'curveball' method for binary networks resulted in `r round(under_z_percent_curveball, 2)`% of networks statistically less nested than null expectations, `r round(no_diff_z_percent_curveball, 2)`% showing no difference, and `r round(over_z_percent_curveball, 2)`% being more nested. The 'quasiswap_count' algorithm for quantitative networks resulted in `r round(under_z_percent_quasiswap, 2)`% of networks being less nested, `r round(no_diff_z_percent_quasiswap, 2)`% showing no difference and `r round(over_z_percent_quasiswap, 2)`% being more nested. Although latitude has a statistically significant impact on residual connectance, it explains only a small portion of the observed variability in residual connectance and nestedness across networks. (connectance: R$^2$ = `r round(connectance_lat_r2, 2)`, _P_ < `r round(connectance_lat_pval, 2)`, **Figure 6c**; NODFc: R$^2$ ~ `r round(normalised_nestedness_lat_r2, 2)`, _P_ = `r round(normalised_nestedness_lat_pval, 2)`, **Figure 6d**). Overall, networks towards higher latitudes showed lower residual connectance and higher nestedness than networks located at lower latitudes. Note that residual connectance and normalised nestedness showed a moderate significant negative correlation (Kendall $\tau$ = `r round(resid_connectance_nestedness_corr_tau, 2)`, _P_ < `r round(resid_connectance_nestedness_corr_pval, 2)`).
